---
title: "covid 19"
author: "Maria Cabanillas"
date: "10/20/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex
$H_0$ There is no difference in females dying compared to males dying  
$H_A$ There is a difference in females dying compared to males dying  


Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent

```{r}
barchartGC(~result + Sex,data=kRdata, main ="Frequency Bar Chart")
```

  
### this bar chart is displaying in total frequency terms our data.  Clearly there are a lot less dead people than alive people as is reflected by the sizes of the bars in the display.  

```{r}
barchartGC(~result + Sex,data=kRdata, type= "percent", main ="percent Bar Chart")
```

The percentage of males with the virus is less than females but the rate of males dying is more than females.   

  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
Of the people that are alive 56 percent are females and 44 percent are males.
Of the people who have died 36 percent are females and 64 percent are males. 

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
The odds of dying if you're a male is 44/1411    
The odds of dying is you're a female is 25/1831  
The odds ratio is (44/1411)/(25/1831)  
The odds of dying is 2.3 to 1  

## Conclusions
There are more men dying than women at a ratio of 2.3 to 1 in Korea during the Covid 19 pandemic.   


